# pml
---
title: "PML"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load all required libraries

```{r}
library(caret)
library(rpart)
library(randomForest)
library(rattle)
```

##Data loading and cleansing 
1. Load csv files
2. Remove columns which have NAs
3. Remove columns which are ids
4. Columns are reduced from 160 to 53
```{r}
training <- read.csv("C:\\pml-training.csv")
testing <- read.csv("C:\\pml-testing.csv")
dim(training)
dim(testing)
NZV <- nearZeroVar(training)
training <- training[,-NZV]
testing <- testing[,-NZV]
dim(training)
dim(testing)
NACols <- sapply(training, function(x) mean(is.na(x))) > 0.95
training <- training[,NACols == FALSE]
testing <- testing[,NACols == FALSE]
dim(training)
dim(testing)
training <- training[,-c(1:5)]
testing <- testing[,-c(1:5)]
dim(training)
dim(testing)
```

##Cross validation
I split the training dataset into training and testing sets with rations 0.7:0.3
```{r}
trainRows <- createDataPartition(y=training$classe, p=0.7, list=FALSE)
trainPart <- training[trainRows,]
testPart <- training[-trainRows,]
dim(trainPart)
dim(testPart)
```

##Model 1 : Use decision tree for building ML model
```{r}
treeModel <- train(classe~., method="rpart", data=trainPart)
fancyRpartPlot(treeModel$finalModel)
treePred <- predict(treeModel, testPart)
confusionMatrix(treePred, testPart$classe)
```
From the confusion matrix, we can see that accuracy is just around 55 percent

##Model 2 : Use random forest model
```{r}
rfModel <- randomForest(classe~., data=trainPart, ntree=500)
treePred2<- predict(rfModel, testPart)
confusionMatrix(treePred2, testPart$classe)
```
##Using random forest model for 20 test cases
```{r}
predict(rfModel, testing)
```


